Predicting and Preventing Employee Turnover through HR Analytics

### 2024-06-07 products/manuscript/Manuscript.qmd
Second update with code.

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

{{\< pagebreak \>}

# Introduction

## General Background Information

@@

## Datasets

**IBM HR Analytics Employee Attrition**

@@

**HR Analytics Case Study**

This data set was also sources from [Kaggle.com](https://www.kaggle.com/datasets/vjchoudhary7/hr-analytics-case-study) and the collaborators are Vijay Choudhary and Aman Kumar. This data set consists of 5 data files and 1 data dictionary file. Each of the data files, contains 4,410 rows and the main file contains 24 variables similar to the IBM file where it shows demographic information, job characteristics, compensation, survey data, etc. This file contains additional information that could help us

## Concerns

The data

*DELETE ----- Describe what the data is, what it contains, where it is from, etc. Eventually this might be part of a methods section.*

## Questions/Hypotheses to be addressed

*State the research questions you plan to answer with this analysis.*

To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above and have the right bibtex key. Then you can include like this:

Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a].

{{< pagebreak >}}

@@

## Schematic of workflow

Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it). @fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender. We store those figures in the `assets` folder.

```{r}
#| label: fig-schematic
#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
#| echo: FALSE
knitr::include_graphics(here("assets","antigen-recognition.png"))
```

## Data aquisition

*As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next.*

## Data import and cleaning

*Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along.*

@@

{{< pagebreak >}}

# Results

*Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project.*

@tbl-summarytable shows a summary of the data.

Note the loading of the data providing a **relative** path using the `../../` notation. (Two dots means a folder up). You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path. You can also use the `here` R package to create paths. See examples of that below. I generally recommend the `here` package.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS("../../results/tables/summarytable.rds")
knitr::kable(resulttable)
```

## Basic statistical analysis

*To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p\<0.05 means statistical significance" interpretation is not valid.*

@fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by gender."
#| echo: FALSE
knitr::include_graphics(here("results","figures","height-weight-stratified.png"))
```

## Full analysis

*Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here.*

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
knitr::kable(resulttable2)
```

{{< pagebreak >}}

# Discussion

## Summary and Interpretation

*Summarize what you did, what you found and what it means.*

## Strengths and Limitations

*Discuss what you perceive as strengths and limitations of your analysis.*

## Conclusions

*What are the main take-home messages?*

*Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end*

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like.

{{< pagebreak >}}

# References

>>>>

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)

hr_data = read.csv("C:/Users/crew_/OneDrive/Escritorio/Guille/UTSA Masters/MSDA/DA6833/PracticumII-Course/guillermogallardo-final-project/data/raw-data/IBM HR Analytics Data.csv", header = TRUE)
```

# Project Overview

@@

# Project Question

We have two main questions that we want to answer during this project.

-   What are the primary factors that contribute to employee attrition?

-   What strategies can be implement to enhance employee retention and improve the overall employee experience?

# Dataset

For this project, I was looking for a dataset with employee data that included a wide range of variables to explore potential relationships with attrition. Below, I will provide an overview of the dataset's origin, author, the types of variables it includes, and other relevant details.

## **IBM HR Analytics Employee Attrition**

@@

## **Variables**

This file includes 35 variables, detailed below, which includes a combination of character and integer types, with some of the character variables being categorical.

```{r Variables}

str(hr_data)
```

## **Gender**

Out of the 1,470 observation in the dataset, 588 are female and 882 are male. This provides a quick snapshot of the gender distribution within the data.

```{r Gender}
table(hr_data$Gender)
```

## **Education**

During data cleaning, we'll use the data dictionary to decode each field. For instance, the education field is labeled from 1 to 5, representing different educational levels: 1 = "Below College," 2 = "College," 3 = "Bachelor's," 4 = "Master's," and 5 = "Doctorate." Notably, about 40% of the dataset holds a Bachelor's degree.

```{r Education}
table(hr_data$Education)
```

@@

@@
